---
title: "Data Manipulation"
output: github_document
date: "2023-10-04"
code_folding: hide
---

```{r}
library(tidyverse)
```

```{r}
litters_df = read.csv("./data/FAS_litters.csv")
litters_df = janitor::clean_names(litters_df)
```


Look at the **dplyr** function

## `select` (selecting variables)

Choose some colomns and not others.

```{r select only `group` and `gd0_weight` column}

select(litters_df, group, gd0_weight)

```

```{r specify a range of colomns}
select(litters_df, group, gd0_weight:gd_of_birth)
```

```{r}
#Specify what we want to get rid of
select(litters_df, -litter_number)
```

Renaming columns ...
*Same for using either code below for renaming*
```{r}
select(litters_df, GROUP = group, LITTer_NUmBer = litter_number)
```

```{r}
rename(litters_df, GROUP = group, LITTer_NUmBer = litter_number)
```

Select helpers

```{r}
# In litters_df, show everything start with "gd"
select(litters_df, starts_with("gd"))
```

```{r}
# move litter_number to the beginning, and keep everything else behind
select(litters_df, litter_number, everything())
```


```{r relocate function in dplyr}
#Does same thing as above (have litter_number in the beginning and everything behind)
relocate(litters_df, litter_number)
```

**Done with selecting variables**


## `filter` (remove rows or keep rows that interested in retaining result)

```{r}
# Only look at gd0_weight less than 22
filter(litters_df, gd0_weight < 22)
```

```{r}
filter(litters_df, gd0_weight >= 22)
```

** `==` is a test for equality, not adding the value into the column (not as `=`)

```{r}
filter(litters_df, gd_of_birth == 20)
```

**Have `!` is to "not equal to"**

```{r eval = FALSE}
# Give all gd_of _birth that not equal to 20
filter(litters_df, !gd_of_birth == 20)
#or
#filter(litters_df, gd_of_birth ! = 20)
```

```{r}
filter(litters_df, gd0_weight >= 22, gd_of_birth == 20)
```

**Non-numerical number**
```{r}
filter(litters_df, group == "Con7")
```
```{r}
filter(litters_df, group == "Mod8")
```

**Join the two groups together (`%in%`)**
```{r combine 2 groups}
# Tell where the group is either of this two groups
filter(litters_df, group %in% c("Con7", "Mod8"))
```
** Done with `filter` function

## `Mutate` function##
### Creating new random variables or modify existing variables###

```{r mutate a new variable "wt_gain"}
mutate(litters_df, wt_gain = gd18_weight - gd0_weight)
```

```{r mutate by adding multiple variables}
mutate(
  litters_df, 
  wt_gain = gd18_weight - gd0_weight,
  group = str_to_lower(group) #modify 'group' variable tp making everything lower case
  )
```

**Done with `mutate` function**

## `arrange` function ##

##### Goal is to make things in order #####




