---
title: "Data Import"
output: github_document
---
Focus on readr.
```{r setup}
library(tidyverse) #the functions exist in tidyverse but not basic R)
library(readxl) #readxl = package contains read_excel function
library(haven) #to read sas file
```
## Read in some data

Read in the litters dataset
```{r}
litters_df = read_csv("./data/FAS_litters.csv")
spec(litters_df)

#follow up step: clean up variable names after importing data
litters_df = janitor::clean_names(litters_df)
```

## Take a look at the data

printing in the console.

```{r}
litters_df
```

```{r}
tail(litters_df) #check the tail of the data to make sure that everything is structured
```

```{r}
skimr::skim(litters_df) #or do it by library(skimr)
```

## Options to read_csv

check out `?read_csv` for more information.

**Example code**
litters_df = read_csv("./data/FAS_litters.csv")
litters_df


##  Other file formats

Read in an excel file.

```{r}
mlb_df = read_excel("./data/mlb11.xlsx", range = "A1:F7")
mlb_df #don't need to clean column names to lower snake case because it's lower letter already
```


Read in a SAS file

```{r}
pulse_df = read_sas("./data/public_pulse_data.sas7bdat")
pulse_df #Need some cleaning up
```

## Comparison with Base R

what about `read.csv` ...?

ALways use underscore `read_csv` instead of dot (`read.csv`)

```{r}
litters_base = read.csv("data/FAS_litters.csv")
litters_readr = read_csv("data/FAS_litters.csv")

litters_base #print the entire dataset by default which give some weird information
litters_readr
```
>Printing difference is the most obvious difference between using "base" and "readr".
"readr" usuaslly faster.

## Export data

Export the mlb sub-table

```{r}
mlb_df #Now is the 6 x 6 small table instead of full dataset
write_csv(mlb_df, "./data/mlb_subtable.csv")
```

